### 2025-10-06 thesis/design/components.tex
Rewrite the design section partially

\subsection{Definitions}\label{design:definitions}

@@

The function utilizes cosine similarity in the embedding space to identify examples
that are semantically closest to the input query $q$, considering both masked natural
language representations and schema distance as described in Section~\ref{design:sampling}.
This approach ensures that examples are weighted by the following properties:

>>>>

\subsection{Functions}\label{design:components}

@@

The function utilizes cosine similarity in the embedding space to identify examples
that are semantically closest to the input query $q$, considering both masked natural
language representations and schema distance as described in Section~\ref{design:embedding}.
This approach ensures that examples are weighted by the following properties:

### 2025-10-06 thesis/design/index.tex
Rewrite the design section partially

The core of the proposed design is the query function $nq$ which utilizes
example selection ($\sigma$), query preprocessing, schema subsetting ($\phi$),
model inference ($\pi$), self-correction ($\rho$) and majority voting ($\nu$)
in order to convert a natural language query $q$ into a corresponding SQL query $\omega$.

% TODO: Revisit
% TODO: Draw diagram of nq, this is an itermediate placeholder
\begin{verbatim}
Query -> Zero-Shot -> Selection -> Generation [ Subsetting -> Few-Shot -> Refinement ] -> Voting
\end{verbatim}
% TODO: Reference figure correctly

\begin{algorithm}

\caption{The $nq$ function}\label{alg:cap}

\begin{algorithmic}[1]
\Require $q \in \mathcal{Q}$, $s \in \mathcal{S}$, $e \in \mathcal{E}$, $v \in \mathcal{V}$, $d \in \mathcal{D}$
\Require $k \in \{ n \in N\ |\ 2n + 1 \}, k \geq 1$
\State $z \gets \pi(\emptyset,q, s)$       \Comment{Zero shot inference}
\State $S \gets \sigma(q,z,s,v,d)$         \Comment{Example selection is performed}
\State $C \gets \emptyset$                 \Comment{The set of query candidates}
\While{$|C| < k$}                          \Comment{While not enough candidates exist}
    \State $s' \gets \phi(C, s)$           \Comment{Identify relevant schema subset}
    \State $c \gets \pi(S, q, s')$         \Comment{Prompt model for next candidate}

    \If{$e(nc)$}                           \Comment{Evaluate the next candidate}
        \State $C \gets \{c\} \cup C$      \Comment{Extend the candidates}
    \EndIf

    \State $c' \gets \rho(q, s', e, v)$  \Comment{Refine via self correction}

    \If{$e(c')$}                          \Comment{Evaluate the refined candidate}
        \State $C \gets \{c'\} \cup C$    \Comment{Extend the candidates}
    \EndIf
\EndWhile
\State \Return $\nu(C, e)$               \Comment{Return most prevalent result}
\end{algorithmic}

\end{algorithm}

This design expands on concepts from DAIL-SQL \citep{DAIL}, X, Y and Z \textcolor{red}{look up}.
The most notable characteristics of \textsc{Natural} lie in it's capabilities for continuous
learning, example selection and portability which allow the system to iteratively
adapt to unseen database schemas.

Prior to using $nq$ sample and database indexing must be done to supply a relevant 
embedding space $\mathcal{V}$ and distance index $\mathcal{D}$.

\subimport{}{sampling}
\subimport{}{definitions}

>>>>

The \textsc{Natural} system implements a multi-stage pipeline that transforms natural
language questions into executable SQL queries through example selection,
schema subsetting, and iterative refinement. The proposed architecture consists
of five core components:

\begin{enumerate}
    \item \textbf{Example Selection} – Identifies semantically and structurally similar
        historical examples using masked embeddings and schema distances.
    \item \textbf{Schema Subsetting} – Reduces database schema complexity by focusing
        on relevant tables and relationships
    \item \textbf{Query Generation} – Few-shot learning with fine tuned models translate natural language queries to SQL
    \item \textbf{Self-Correction} – Iteratively refines queries through execution feedback and error analysis
    \item \textbf{Consensus Voting} – Selects the most reliable result from multiple
        generation attempts
\end{enumerate}

The system processes queries through the following pipeline:

% TODO: Make a nice diagram for this
\begin{verbatim}
  Natural Language Query 
→ Sketch Generation
→ Example Selection
→ k-times [
    → Schema Subsetting
    → Few-Shot Generation
    → Self-Correction
  ]
→ Consensus Voting
→ SQL Query
\end{verbatim}

This design builds upon concepts from DAIL-SQL \citep{DAIL} and incorporates novel
contributions in schema-aware example selection and graph-based structural similarity.

\subimport{}{initialization}

\subimport{}{components}

\subimport{}{composition}
